---
title: "R Notebook"
output: html_notebook
---

CMTH642 Data Analytics: Advanced Methods
Lab 9 - Solutions

```{r}
install.packages("class")
install.packages("gmodels")
library(class)
library(gmodels)

```
1. Download Prostate Cancer.csv from the course shell. The dataset contains patients
who have been diagnosed with either Malignant (M) or Benign (B) cancer.
```{r}
 prc <- read.csv(file="C:\\Users\\Mohsen\\Desktop\\Prostate_Cancer.csv", stringsAsFactors = FALSE)

```
2. Check the data characteristics. Remove the first variable (id) from the dataset. Get the
numbers of patients in each class.
```{r}
str(prc)
```
```{r}
prc <- prc[-1]

```

```{r}
table(prc$diagnosis_result)
```

```{r}
prc$diagnosis <- factor(prc$diagnosis_result,
levels = c("B", "M"),
labels = c("Benign", "Malignant"))
round(prop.table(table(prc$diagnosis)) * 100, digits = 1)
```
3. Create a function to normalize a variable. Normalize the numeric variables in the dataset
to transfer their values to a common scale
```{r}
normalize <- function(x) {
return ((x - min(x)) / (max(x) - min(x))) }
```

```{r}
prc_n <- as.data.frame(lapply(prc[2:9], normalize))
prc_n <- cbind(prc$diagnosis, prc_n)
prc_n
```

```{r}
summary(prc_n$radius)
```
4. Split the dataset into training and test sets
```{r}
set.seed(123)
index <- sample(1:nrow(prc_n), 0.65 *nrow(prc_n))
prc_train <- prc_n[index,]
prc_test <- prc_n[-index,]
```
5. Use the KNN algorithm to predict the cancer.
```{r}
prc_train_labels <- prc_train[,1]
prc_test_labels <- prc_test[,1]

```

```{r}
prc_test_pred <- knn(train = prc_train[,2:9],
test = prc_test[,2:9],
cl = prc_train[,1], k=10)

```
6. Evaluate the model performance
```{r}
table(Actual=prc_test_labels, Predicted=prc_test_pred)
```

